---
title: "Geostatistics"
output:
  github_document:
    pandoc_args: --webtex

---
# Geostatistics

## Check Working directory


```{r include = FALSE}
getwd()
list.files()
getwd()


```

## R packages
```{r include = TRUE}
library(geoR)
library(ggplot2)
# library(akima)
# library(fields)
# require(graphics)
# rmarkdown::render("Ejemplo_aquifer_geoR.Rmd")
# rmarkdown::render("Example.Rmd")
```
- [geoR documentation](https://cran.r-project.org/web/packages/geoR/geoR.pdf)


## Data

```{r echo = T, results = "hide"}
aquifer <- read.table("aquifer.txt", head = TRUE, dec = ",")
head(aquifer)
summary(aquifer)
```

Summary aquifer
```{r echo = FALSE}
knitr::kable(summary(aquifer), label = "Summary")
```
## GEO_Data
### Data frame to GEO_Data

- Documentation [as.geodata](https://cran.r-project.org/web/packages/geoR/geoR.pdf#page=5)

```{r warning = FALSE}
aquiferg <- as.geodata(aquifer)
summary(aquiferg)
```
### Plot GEO_Data Objtect
- Documentation [plotgeodata](https://cran.r-project.org/web/packages/geoR/geoR.pdf#page=85)


```{r warning = FALSE}
plot(aquiferg, qt.col = c("purple",
                         "pink",
                         "green",
                         "yellow"))
``` 


Scatter 3d
```{r warning = FALSE}
plot(aquiferg, scatter3d = T)
``` 

the mean part of the model

```{r warning = FALSE}
plot(aquiferg, trend = "1st")
plot(aquiferg, trend = "2nd")

``` 
# Mean part of the model

## Model 1
```{r echo = T , results = "hide"}

reg1 <- lm(Profundidad ~ Este + Norte, data = aquifer)
residuales1  <-  residuals(reg1)

summary(reg1)
anova(reg1)

```

```{r echo = F} 
pander::pander(summary(reg1))
pander::pander(anova(reg1))
```


## Model 2

```{r echo = T, results = "hide" }

reg2 <- lm(Profundidad ~ Este + Norte +
           I(Este^2) + I(Norte^2) +
           I(Este * Norte),
           data = aquifer)
residuales2  <-  residuals(reg2)
summary(reg2)
anova(reg2)

```

```{r echo = F}

pander::pander(summary(reg1))
pander::pander(anova(reg1))

```

## Model 3

```{r echo = T, results = "hide" }

reg3 <- lm(Profundidad ~ Este * Norte,
           data = aquifer)
residuales3  <-  residuals(reg3)
summary(reg3)
anova(reg3)

```

```{r echo = F}

pander::pander(summary(reg3))
pander::pander(anova(reg3))

```
# Empirical semivariogram

- Documentation [variog](https://cran.r-project.org/web/packages/geoR/geoR.pdf#page=133)


```{r echo = T, results = "hide" }

vari2 <- variog(aquiferg, trend = "1st")
vari2Cloud <- variog(aquiferg, op = "cloud", trend = "1st")
vari2BinCloud <- variog(aquiferg,
                       max.dist = 200,
                       op = "cloud",
                       bin.cloud = TRUE)
vari2Sm <- variog(aquiferg,
                  trend = "1st",
                  op = "sm",
                  band=11)

```

```{r echo = T}
par(mfrow = c(2, 2), mar = c(3, 3, 1, 1), mgp = c(2, 1, 0))
     plot(vari2, main = "binned variogram")
     plot(vari2Cloud, main = "variogram cloud")
     plot(vari2BinCloud,main = "clouds for binned variogram")
     plot(vari2Sm, main = "smoothed variogram")
```

## Classic estimation

```{r echo = T, results  = "hide"}

vari1 <- variog(aquiferg)
vari2 <- variog(aquiferg, trend = "1st")
vari3 <- variog(aquiferg, trend = "2nd")

```
```{r echo = T}
plot(vari1, main =  "Constant")
plot(vari2, main  = "Trend 1 ")
plot(vari3, main  = "Trend 2 ")
```



## Anisotropy

```{r echo = T, results = "hide"}
vari_0 <- variog(aquiferg,
                 trend = "1st",
                 max.dist = 200,
                 dir = 0)

vari_45 <- variog(aquiferg,
                  trend = "1st",
                  max.dist = 200,
                  dir = pi / 4)
vari_90 <- variog(aquiferg,
                  trend = "1st",
                  max.dist = 200,
                  dir = pi / 2)
vari_135 <- variog(aquiferg,
                   trend = "1st",
                   max.dist = 200,
                   dir = 3 * pi / 4)
par(mfrow = c(2, 2),
    mar = c(3, 3, 1, 1),
    mgp = c(2, 1, 0))

plot(vari_0, main = "vari 0")
plot(vari_45, main = "vari 45")
plot(vari_90, main = "vari 90")
plot(vari_135, main = "vari 195")

```

## Theoretical estimation

- Documentation [eyefit](https://cran.r-project.org/web/packages/geoR/geoR.pdf#page=25)
- Documentation [variofit](https://cran.r-project.org/web/packages/geoR/geoR.pdf#page=129)
- Documentation [likfit](https://cran.r-project.org/web/packages/geoR/geoR.pdf#page=60)



```{r echo = T, results = "hide"} 

var1 <- variog(aquiferg,
               trend = "1st",
               max.dist = 200)


#ini1 <- eyefit(var1)
#cov.model  sigmasq phi   tausq kappa kappa2   practicalRange
#1      wave 30805.52  13 8984.94  <NA>   <NA> 38.8889336320589
ini1 <- c(30805.52, 13)
fitvar1 <- variofit(var1,
                    cov.model = "wave",
                    ini1,
                    fix.nugget = TRUE,
                    nugget = 8984.94,
                    wei = "equal")

fitvar2 <- variofit(var1,
                    cov.model = "wave",
                    ini1,
                    fix.nugget = TRUE,
                    nugget = 8984.94,
                    wei = "npairs")

fitvar3 <- variofit(var1,
                    ini1,
                    fix.nugget = TRUE,
                    nugget = 8984.94,
                    wei = "cressie")


fitvar4 <- likfit(aquiferg,
                  coords = aquiferg$coords,
                  data = aquiferg$data,
                  trend = "1st",
                  ini.cov.pars = ini1,
                  fix.nugget = T,
                  nugget = 8984.94,
                  cov.model = "wave",
                  lik.method = "ML")

fitvar5 <- likfit(aquiferg,
                  coords = aquiferg$coords,
                  data = aquiferg$data,
                  trend = "1st",
                  ini.cov.pars = ini1,
                  fix.nugget = T,
                  nugget = 8984.94,
                  cov.model = "wave",
                  lik.method = "REML")
```


```{r echo = T}
plot(var1,
     xlab = "h",
     ylab = "semivariance",
     cex.lab = 1.3,
     cex.axis = 1.2,
     main = "Semivariogram Estimation",
     col.main = 4, cex.main = 1.3)
lines(fitvar1, col = 1)
lines(fitvar2, col = 2)
lines(fitvar3, col = 3)
lines(fitvar4, col = 4)
lines(fitvar5, col = 5)
legend(130, 18000,
       c("MCO", "MCPnpairs", "MCPcressie", "ML", "REML"),
       lwd = 2,
       col = 1:5,
       box.col = 9,
       text.col = 1:5)
```

# Results 

```{r echo = T}
summary(fitvar1)
summary(fitvar2)
summary(fitvar3)
summary(fitvar4)
summary(fitvar5)

```

# Prediction

```{r echo = T}
pred.grid <-  expand.grid(seq(-145, 112, l = 100),
                          seq(9, 184, l = 100))
# kriging calculations
kc_pred <- krige.conv(aquiferg,
                 loc = pred.grid,
                 krige = krige.control(obj.model = fitvar1))


# displaying predicted values
```

```{r echo = T}

predicted_df <- data.frame(X = pred.grid$Var1,
                           Y = pred.grid$Var2,
                           Pred = kc_pred$predict,
                           Var = kc_pred$krige.var)

coors <- data.frame(X = aquiferg$c[, 1],
                    Y = aquiferg$c[, 2])
plot_prediction <-
    ggplot() +
        geom_tile(data = predicted_df,
                  mapping = aes(x = X, y = Y, fill = Pred)) +
        scale_fill_viridis_c(option = "inferno",
                             direction = -1) +
        geom_point(data = coors, mapping = aes(x = X, y = Y), shape = 4)

plot_variance <-
    ggplot() +
        geom_tile(data = predicted_df,
                  mapping = aes(x = X, y = Y, fill = Var)) +
        scale_fill_viridis_c(option = "inferno",
                             direction = -1) +
        geom_point(data = coors, mapping = aes(x = X, y = Y), shape = 4) +
        theme_void()

plot_prediction
plot_variance
```




